Improved Bounds for Online Balanced Repartitioning

### 2020-04-06 paper/waoa20clustering.tex
Cascade sketch

  		 \item $O(\ell)$-competitive algorithm for $k=4$ and $k=5$

@@

\subsection{Cascade}

@@

\begin{theorem}
  A single repartition of DET2 migrates nodes among at most $p(k)^2+(k+1)$ clusters, where $p(k)$ is the number of integer partitions $k$.
  The corollary is, the cost of the repartition is bounded by $O(k \cdot \alpha \cdot e^{c\cdot \sqrt{k}})$ for a constant $c$.
\end{theorem}

\begin{proof}
  \maciek{Danger zone, nothing is finished here}

  A \emph{signature} of a cluster $C$, denoted $S(C)$ is a multiset of sizes of components located in this cluster.
  Note that there are $p(k)$ possible signatures.

  Fix a repartition.

  Consider the repartition from the perspective of a single cluster.
  Prior to the repartition, it had a signature $S^-(C)$, and after the repartition it has a signature $S^+(C)$.

  Now we construct a graph $G$ of component exchange among clusters.
  The vertices are the clusters.
  If a pair of clusters $C_1, C_2$ exchange components during the repartition, we produce an edge and we label it with the \emph{signature of exchange}, i.e., a pair of cluster signatures (both ways), limited to the exchanged clusters between $C_1,C_2$ only (i.e., not necessarily summing to $k$, and  note that also not necessarily having the same size!).

  Proof by contradiction.
  Assume that there exists a component in $G$ of size greater than $p(k)^2+(k+1)$.

\end{proof}

>>>>

@@

\subsection{Cascade bounds and an improved algorithm for general k, l}

@@

Now we bound the cost of a single repartition of DET2 independent of $\ell$.
First, we bound the cycle length in one of optimal solutions.

\begin{theorem}
  \label{th:cascade-cycles}
  Consider a single repartition of DET2 caused by node merge.
  There exists a sequence of node migrations that realizes this repartition, where each node-disjoint cycle has length at most $k$.
\end{theorem}
Describe the graph. Edges and labels.

Proof by contradiction.
We call a sequence of node migrations that realize the repartition \emph{a solution to repartition}.
Assume that each cost-optimal solution to repartition contains a cycle of length at least $k+1$.

Fix a solution to repartition that has the shortest length of the longest cycle, and has the least edge-disjoint cycles of such length.

\begin{proof}
\end{proof}

\begin{theorem}
  Consider a single repartition of DET2.
  The number of clusters it involves is bounded by $k^k$.
  \label{th:cascade}
\end{theorem}

\begin{proof}
  Each node has degree at most $k$. Combine with Theorem \ref{th:cascade-cycles}.

Important technical details:
\begin{itemize}
  \item There are no cycles attached to another cycles by just one vertex (by equivalent signature skipping). This is important, as otherwise we just have a bound on cycles, but the graph contains a chain of cycles.
  Another formulation (Stefan's): each edge is a part of a cycle that goes through the "origin" part of the graph.
\end{itemize}
\end{proof}

\begin{theorem}
  Algorithm DET2 is $O(\ell \cdot k\cdot k^k)$-competitive.
\end{theorem}

\begin{proof}
  Use Theorem~\ref{th:cascade}. Then, linearize the inner-component communication using the technique from SIDMA.
\end{proof}
